---
title: "Find stable topics in STM with Jaccard similarity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{find_stable_topics-jaccard}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

#mods <- readRDS("../mod.RDS")
```

This vignette shows how to use `topicl` to find stable (reproducible)
topics with Jaccard similarity. This method is presented in (Koltsov
et al., 2018), but can be shortly summarized as follows.

A stable topic is a topic that persists across fits of a topic model to
the data. Thus, keeping model parameters fixed, a stable topic must
appear in a majority of fits. Across fits, we compare topics to each
other to test whether a given pair is the same. For comparison, we use
compute Jaccard similarity on N top-words for a topic. Papers by
Koltosov and colleagues advise fitting at least five models and
consider a given pair of topics identical if their Jaccard similarity
is =< 90%. Then, if a stable topic persisted across at least 3 out of 5
model runs, you can consider this topic stable.

In this vignette, we will fit five STM models to `poliblog5k` (see
`?stm::poliblog5k` for details) and consider how many topics persists
using `topicl`.

```{r, setup, message=F, warning=F}
library(topicl)
library(stm)
library(tictoc)
library(furrr)
```

`fit_stm` specifies STM parameters. We increase the maximum number of
iterations for the EM algorithm and ensure that each iteration is
performed with `max_it = 1000` and `emtol = 0` - see `?stm` for
details. In my experience, with an increased number of iterations on these
data, topic similarity across fits also increases. I cannot say if `K
= 25` is optimal for these data, but this number of topics is used in
STM documentation.

With <random.org>, we generated five random `seeds` for each model.

```{r model-specification}
fit_stm <- function(seed) {
  K = 25
  max_it = 1000
  emtol = 0
  
  mod <- stm(poliblog5k.docs, 
            poliblog5k.voc, K=K,            
            data=poliblog5k.meta,
            max.em.its=max_it,
            emtol = emtol,
            init.type="Random",
            seed = seed,
            verbose = F)
  return(mod)
}


seeds <- c(9934,9576,1563,3379,8505)
```

With `purrr::future_map()`, we fit five models with different seeds in
parallel. Given the high number of iterations, it can take up to 40
minutes. 

```{r, fit-models}
tic()
plan(multisession, workers = 5) # toggle multiple parallel processes
mods <- future_map(seeds, fit_stm,                   
                   .options = furrr_options(seed=T))
plan(sequential) # toggle back to sequential
toc()
```

We use `topicl::compare_solutions()` to calculate topic similarity on
the level of `depth = 10` top terms for a topic. By default, top terms
are produced by ranking term probabilities. Please note that (Koltsov
et al., 2018) advice using 100 top terms.

```{r compare-topics}
results <- compare_solutions(mods, depth = 10)
```

What topics are similar and appear across fits is helpful to visualize
with a network plot. Given output of `topicl::compare_solutions()` and a
threshold, `topicl::viz_comparisons()` will plot a topic-topic network.
Here, we plot connections between topics =< 80% similarity.

```{r visualize-topic-similarity}
viz_comparisons(results, 0.8)
```

With `topic::filter_topics()`, we can get a table with topic IDs and
counts of their distinct pairs in other models.

```{r count-topic-appearance}
filter_topics(results, 0.5, 2)
```

# References

1. Koltsov, S., Pashakhin, S., & Dokuka, S. (2018). A Full-Cycle
   Methodology for News Topic Modeling and User Feedback Research. In
   S. Staab, O. Koltsova, & D. I. Ignatov (Eds.), Social Informatics
   (Vol. 11185, pp. 308–321). Springer International Publishing.
   <https://doi.org/10.1007/978-3-030-01129-1_19>


# Appendix

```{r sessionInfo}
sessionInfo()
```
